---
title: "Lecture 2 - 9-6"
output: html_document
---

# to load tidyverse package: 

```{r}
library(tidyverse)
```

# Basic Computations & Arithmetic
```{r}
2 + 3
(18/3 + 1*2)^(4 - 2) - 6.2

x = (18/3 + 1*2) ^ (4 - 2) - 6.18273
y = c(1, 3, 6, 9)
x + y
```

# Apply functions to objects

```{r}
x = runif(20) #runinf = selects a random number from a uniform distribution
x

mean(x)
var(x)
```

# create a data frame (tibble) then create different subset variable names with a length of four

```{r}
example_df = tibble(
  vec_numeric = 5:8, 
  vec_char = c("My", "name", "is", "Jeff"), 
  vec_logical = c(TRUE, TRUE, TRUE, FALSE),
  vec_factor = factor(c("male", "male", "female", "female"))
)
```

# examine classes to see if they are character, integer, factors, etc.

```{r}
class(example_df$vec_numeric) # this is an integer
class(example_df$vec_factor) # this is a factor/categorical variable
```

# Basic Plots

```{r}
set.seed(1234) #set seed to ensure reproducibility

plot_df = tibble( #define data frame that contains x and y
  x = rnorm(1000, sd = .5),
  y = 1 + 2*x + rnorm(1000)
)

ggplot(plot_df, aes(x = x)) + geom_histogram() #create histogram of x. Can use 'binwidth' to pick diff width values for the bars

ggplot(plot_df, aes(x = x, y = y)) + geom_point() #create scatterplot of x and y
```

# to clear workspace, which enables you to restart and double chek that your code will re-run properly
```{r}
rm(list = ls())
```
# to check working directory
```{r}
getwd()
```
# to plot and save a scatterplot
```{r}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```
